\textbf{How to measure maximally precise radial velocities from a set of spectra:\\ The time-independent spectrum case}

### 2019-08-26 paper/eprv.tex
note to self
clarity

For our purposes, the \CRLB\ is that the \emph{inverse variance} $\sigma_v^{-2}$ of any velocity
estimator (that is, the information in the velocity estimator) must not be larger than the
second derivative of the \emph{negative} log likelihood function:
\begin{equation}
\frac{1}{\sigma_v^2} \leq E\left[\frac{\dd^2}{\dd v^2}(-\ln L)\right] \quad,
\end{equation}
where $L$ is the likelihood, or the probability of the data given the velocity,
and the expectation $E[\cdot]$ is taken over all possible data (that is, it is an
expectation under the likelihood).
This expectation value of a second derivative is the Fisher Information
(information has units of inverse variance).
This is the information-theory object that is frequently used in
cosmological experiment forecasts (\todo{CITE SHIT}),
and it is the quantity used for what we often call the ``formal error'' or
``formal uncertainty'' on a least-square fit (\todo{CITE fitting a line}).
The bound is an upper bound on the inverse variance, and thus a lower bound
on the uncertainty.

If there are many nuisance parameters, a Bayesian can marginalize them
out to make $L$ the marginalized likelihood for the velocity, or a frequentist
can include them in the estimation and compute an inverse covariance matrix $C^{-1}$
for the full parameter vector.
This will then will have the bound
\begin{equation}
v^T\cdot C^{-1}\cdot v \leq v^T\cdot Q\cdot v
\end{equation}
\begin{equation}
Q_{ij} \equiv E\left[\frac{\dd^2}{\dd\theta_i\,\dd\theta_j}(-\ln L)\right] \quad,
\end{equation}
where $Q$ is the information tensor, $i,j$ are indices into that
tensor, corresponding parameters are $\theta_i, \theta_j$, and $v$ is
any arbitrary vector.
That is, the bound is not just on every parameter, but really on on
any combination of parameters.

@@

The most optimistic assumptions we can possibly make are the following.
\begin{itemize}
\item The spectrograph is perfectly calibrated in a wavelength sense.
\item The spectrograph is consistently and repeatably calibrated in a
  flux sense. That is, the spectrum does not need to be perfectly
  continuum-normalized or perfectly flux-normalized, but it is
  calibrated near-perfectly \emph{consistently} across epochs (exposures).
  We will break this assumption in our experiments, below.
\item There are no residual telluric absorptions nor any residual sky
  emission in the spectra. That is, the sky is near-perfectly calibrated
  and removed.
  We will also break this assumption in our experiments, below.
\item The \emph{true} spectrum (that is, the latent spectrum that
  would be observed if the data were far, far better) of the star is
  known at better accuracy and precision than the main noise sources.
  We will break this assumption too.
  The whole point of this paper is to ask whether we can create a
  good-enough proxy for the true spectrum.
\item There is no time dependence (no epoch-to-epoch variations) in the
  stellar spectrum.
\item All noise sources contributing to the spectral data---which
  include at a minimum photon noise, read noise, residual telluric
  issues, residual wavelength-calibration issues, and residual stellar
  spectral variability---are, when summed together, indistinguishable
  from being zero-mean, Gaussian, and additive.
\item The complete variance tensor of the total noise on the spectrum is known, and
  known correctly. That is, the variance of the noise at every pixel
  is known, as is the covariance between nearby pixels, and in the
  same \emph{true} sense that the true spectrum of the star is known.
\end{itemize}
It is worthy of note that every formula (that we know) in the
literature for maximum \RV\ precision (\todo{cite Butler, Bouchy,
  Figueira, Lovis}) makes all of these same assumptions,
plus additional \emph{even more restrictive assumptions}.
Under these disturbingly unrealistic assumptions, the \CRLB\ becomes
\begin{equation}\label{eq:crlb}
\frac{1}{\sigma_v^2} \leq \left[\frac{\dd f}{\dd v}\right]\T\cdot C^{-1}\cdot\left[\frac{\dd f}{\dd v}\right]
\end{equation}
where the derivatives are of the true spectrum with respect to
velocity (which can be thought of as the derivative of a Doppler
operator with respect to velocity acting on the true spectrum), the
derivatives are column vectors, and $C^{-1}$ is the inverse of the covariance matrix
describing the true variance and covariances of the noise process.

@@

>>>>

For our purposes, the \CRLB\ is that the \emph{inverse variance} $\sigma_v^{-2}$ of any velocity
estimator (that is, the information in the velocity estimator) must not be larger than the
second derivative of the \emph{negative} log likelihood function:
\begin{equation}
\frac{1}{\sigma_v^2} \leq E\left[\frac{\dd^2}{\dd v^2}(-\ln L)\right] \quad,
\end{equation}
where $L$ is the likelihood, or the probability of the data given the radial velocity $v$,
and the expectation $E[\cdot]$ is taken over all possible data (that is, it is an
expectation under the likelihood).
This expectation value of a second derivative is the Fisher Information
(information has units of inverse variance).
This is the information-theory object that is frequently used in
cosmological experiment forecasts (\todo{CITE SHIT}),
and it is the quantity used for what we often call the ``formal error'' or
``formal uncertainty'' on a least-square fit (\todo{CITE fitting a line}).
The bound is an upper bound on the inverse variance, and thus a lower bound
on the uncertainty.

If there are many nuisance parameters, a Bayesian can marginalize them
out to make $L$ the marginalized likelihood for the velocity, or a frequentist
can include them in the estimation and compute an inverse covariance matrix $C^{-1}$
for the full parameter vector.
This will then will have the bound
\begin{equation}
u^T\cdot C^{-1}\cdot u \leq u^T\cdot Q\cdot u
\end{equation}
\begin{equation}
Q_{ij} \equiv E\left[\frac{\dd^2}{\dd\theta_i\,\dd\theta_j}(-\ln L)\right] \quad,
\end{equation}
where $Q$ is the information tensor, $i,j$ are indices into that
tensor, corresponding parameters are $\theta_i, \theta_j$, and $u$ is
any arbitrary vector.
That is, the bound is not just on every parameter, but really on on
any combination of parameters.

@@

The most optimistic assumptions we can possibly make are the following.
\begin{itemize}
\item The spectrograph is perfectly calibrated in a wavelength sense.
\item The spectrograph is consistently and repeatably calibrated in a
  flux sense. That is, the spectrum does not need to be perfectly
  continuum-normalized or perfectly flux-normalized, but it is
  calibrated near-perfectly \emph{consistently} across epochs (exposures).
  We will break this assumption in our experiments, below.
\item There are no residual telluric absorptions nor any residual sky
  emission in the spectra. That is, the sky is near-perfectly calibrated
  and removed.
  We will also break this assumption in our experiments, below.
\item The \emph{true} spectrum (that is, the latent spectrum that
  would be observed if the data were far, far better) of the star is
  known at better accuracy and precision than the main noise sources.
  We will break this assumption too.
  The whole point of this paper is to ask whether we can create a
  good-enough proxy for the true spectrum.
\item There is no time dependence (no epoch-to-epoch variations) in the
  stellar spectrum.
\item All noise sources contributing to the spectral data---which
  include at a minimum photon noise, read noise, residual telluric
  issues, residual wavelength-calibration issues, and residual stellar
  spectral variability---are, when summed together, indistinguishable
  from being zero-mean, Gaussian, and additive.
\item The complete variance tensor of the total noise on the spectrum is known, and
  known correctly. That is, the variance of the noise at every pixel
  is known, as is the covariance between nearby pixels, and in the
  same \emph{true} sense that the true spectrum of the star is known.
\end{itemize}
It is worthy of note that every formula (that we know) in the
literature for maximum \RV\ precision (\todo{cite Butler, Bouchy,
  Figueira, Lovis}) makes all of these same assumptions,
plus additional \emph{even more restrictive assumptions}.
Under these disturbingly unrealistic assumptions, the \CRLB\ becomes
\begin{equation}\label{eq:crlb}
\frac{1}{\sigma_v^2} \leq \left[\frac{\dd f}{\dd v}\right]\T\cdot C^{-1}\cdot\left[\frac{\dd f}{\dd v}\right]
\end{equation}
where the derivatives are of the true spectrum $f$ with respect to
velocity $v$ (which can, in turn, be thought of as the derivative of a Doppler
operator with respect to velocity acting on the true spectrum; CITE WOBBLE), the
derivatives are column vectors, and $C^{-1}$ is the inverse of the covariance matrix
describing the true variance and covariances of the noise process.

@@

HOGG: OPEN by giving the LF and what it's optimum is, in the simplest case.
Parameterize $f(v)$ where $v$ is the \RV.
